---
title: "vagalumeR: Access the Vagalume API via R"
output: html_document
---

This package allows you to access the Vagalume API via R. You can get
information about specific artists/bands, as their related artists,
and information about the lyrics, as the top lyrics of an artist/band
and the text of a song itself. 


The `vagalumeR` package is developed under control version using Git 
and is hosted in GitHub. You can download it from GitHub using `devtools`:
```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("brunaw/vagalumeR")
```

For exploring the content of the package, you may try:
```{r, eval=FALSE}
library(vagalumeR)
packageVersion("vagalumeR")
ls("package:vagalumeR")
help(package = "vagalumeR")
```

------------------------------------------------------------------------

Some examples of the usage of this package are above:
```{r, echo = FALSE, message=FALSE}
library(vagalumeR)
```

```{r, message=FALSE}
artists <- c("the-beatles", "madonna", "chico-buarque",
             "the-rolling-stones", "molejo")

library(tidyverse)
library(plyr)

ldply(map(artists, artistInfo), data.frame)
```

Where:

  - `id` is the identifier of the artist/band;
  - `name` is the name of  the artist/band;
  - `views` is the number of views received by the artist/band at the
  Vagalume Website;
  - `pos` is the position of the artist/band at the Vagalume Ranking;
  - `period` is the period of evaluation;
  - `uniques` counts the number of unique views of the artist/band;
  - `points` is the pontuation of the artist/band at the Vagalume 
  Ranking;

```{r, message=FALSE}
disc <- ldply(map(artists, discography), data.frame)

disc[sample(nrow(disc), 10), ]
xtabs(~name, disc)
```

Where:

  - `album.id` is the identifier of album;
  - `album.name` is the name of  the album;
  - `label` is the record label of the album;
  - `date` is the of release of the album;

```{r, message=FALSE}
rel <- ldply(map(artists, relatedInfo), data.frame)

rel[sample(nrow(rel), 10), ]
```
Where:

  - `rel.id` is the identifier of the related band/artist;
  - `related` is the name of the related band/artist;


```{r, message=FALSE}
#---------
net <- data.frame(Source = rel$name, 
                  Target = rel$related)
networkD3::simpleNetwork(net,
                         fontSize = 14,
                         linkColour = "turquoise",
                         nodeColour = "tomato")
```

```{r, message=FALSE}
songs <- ldply(map(artists, songNames), data.frame)

songs[sample(nrow(songs), 10), ]
```

Where:

  - `song.id` is the identifier of the song;
  - `song` is the name of the song;

```{r, message=FALSE}
# lyr <- map(songs$song.id[1:5], lyrics, type = "id")
songs$song.id[1:5]
# lyr[sample(nrow(lyr), 2), ]
```

Where:

  - `language` is the language of the song;
  - `song` is the name of the song;
  - `language` is the language of the song (1 = portuguese,
  2 = english)
  - `text` is the lyrics text of the song;
  - `tr.text` is the translation text of the song;

```{r}
# library(tm)
# # Creates the Corpus
# m <- lyr[lyr$name == "The Beatles", ]
# 
# cps <- VCorpus(VectorSource(m$text),
#                  readerControl = list(language = "en"))
# 
# cps <- tm_map(cps, FUN = content_transformer(tolower))
# cps <- tm_map(cps, FUN = removePunctuation)
# cps <- tm_map(cps, FUN = removeNumbers)
# cps <- tm_map(cps, FUN = stripWhitespace)  
# cps <- tm_map(cps,
#                 FUN = removeWords,
#                 words = stopwords("english"))
# dtm <- DocumentTermMatrix(cps)
# 
# frq <- slam::colapply_simple_triplet_matrix(dtm, FUN = sum)
# frq <- sort(frq, decreasing = TRUE)
# 
# library(lattice)
# # Shows the most common words in songs from The Beatles
# barchart(head(frq, n = 45), xlim = c(0, NA),
#          col =  "violet", 
#          xlab = "Frequência",
#          ylab = "Bi-gramas",
#          strip = strip.custom(bg = "white"))
```


